Covariate analysis of UFL-Mayo-ISB RNASeq from cerebellum

### 2016-01-29 code/Rmd/Mayo_Cerebellum_RNASeq.Rmd
Minor changes

# Initial design matrix with covariates under consideration
DM = getDesignMatrix(COVARIATES, Intercept = F)
```
Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Diagnosis), shape=as.factor(Source), size=as.numeric(AgeAtDeath)))
p <- p + theme_bw() + theme(legend.position="top") 
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p

# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(VOOM.GENE_EXPRESSION$E)), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 200, minSize=3)

@@

### Clustering residual gene expression
```{r decompse.normalise.data1, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION[,-(1)], scale.=T)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Diagnosis), shape=as.factor(Source), size=as.numeric(AgeAtDeath)))
p <- p + theme_bw() + theme(legend.position="top") 
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p

# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(RESIDUAL.GENE_EXPRESSION[,-(1)])), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 200, minSize=3)

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp}
# Fit contrast
contrast = makeContrasts(contrasts=c("DiagnosisControl-DiagnosisAD",
                                     "DiagnosisControl-DiagnosisPSP",
                                     "DiagnosisControl-DiagnosisPathologic_Aging",
                                     "DiagnosisPathologic_Aging-DiagnosisPSP",
                                     "DiagnosisPathologic_Aging-DiagnosisAD",
                                     "DiagnosisPSP-DiagnosisAD"), 
                         levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT.CONTR = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)

>>>>

# Initial design matrix with covariates under consideration
DM = getDesignMatrix(COVARIATES, Intercept = F)
```
Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data1, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Diagnosis), shape=as.factor(Source), size=as.numeric(AgeAtDeath)))
p <- p + theme_bw() + theme(legend.position="top") 
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p
```
```{r decompse.normalise.data2, fig.height=10, fig.width=25, results='asis'}
# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(VOOM.GENE_EXPRESSION$E)), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 200, minSize=3)

@@

### Clustering residual gene expression
```{r decompse.normalise.data3, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION[,-(1)], scale.=T)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Diagnosis), shape=as.factor(Source), size=as.numeric(AgeAtDeath)))
p <- p + theme_bw() + theme(legend.position="top") 
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)
p
```
```{r decompse.normalise.data4, fig.height=8, fig.width=25, results='asis'}
# Tree based clustering
sampleTrees = flashClust::hclust(dist(t(RESIDUAL.GENE_EXPRESSION[,-(1)])), method = "average")
tree = WGCNA::cutreeStatic(sampleTrees, cutHeight = 200, minSize=3)

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = RESIDUAL.GENE_EXPRESSION$ensembl_gene_id,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(RESIDUAL.GENE_EXPRESSION, Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=16}
# Fit contrast
contrast = makeContrasts(contrasts=c("DiagnosisControl-DiagnosisAD",
                                     "DiagnosisControl-DiagnosisPSP",
                                     "DiagnosisControl-DiagnosisPathologic_Aging",
                                     "DiagnosisPathologic_Aging-DiagnosisPSP",
                                     "DiagnosisPathologic_Aging-DiagnosisAD",
                                     "DiagnosisPSP-DiagnosisAD"), 
                         levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT.CONTR = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT.CONTR = eBayes(FIT.CONTR)
